---
title: "Structural Bioinformatics (pt.1)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics

Download a CSV file from the PDB site (accessible from “Analyze” > “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and use it to answer the following questions:

Q: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.

```{r}
stats <- read.csv("Data Export Summary.csv", row.names = 1)
stats

```


```{r}
ans <- round(stats$Total / sum(stats$Total) * 100 , 2)
names(ans) <- rownames(stats)
ans
```


Q: What proportion of structures in the PDB are protein?

```{r}
round(sum(stats$Proteins) / sum(stats$Total) * 100, 2)
```


Q: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?


Bio3D is an R package for structural bioinformatics. Features include the ability to read, write and analyze biomolecular structure, sequence and dynamic trajectory data.


```{r}
library(bio3d)

pdb <- read.pdb("1hsg")
pdb
```

Note that the attributes (+ attr:) of this object are listed on the last couple of lines. To find the attributes of any such object you can use:

```{r}
attributes(pdb)
```






