---
title: "Class14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Read our data for today


```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)

nrow(counts)
```

We will compare the control vs. treated count data contained in `counts` object. 

We will first extract the control columns by looking up their colnames in the `metadata` object.

Accces just the "control" metadata, then take the mean value for each gene (i.e. each row)

```{r}
control <- metadata[metadata[,"dex"]=="control",]
control.mean <- rowSums( counts[ ,control$id] )/4 
names(control.mean) <- counts$ensgene
```

Repeat the same process for `treated`.

```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated.mean <- rowSums(counts[,treated$id])/4
treated.mean
```


We will combine our meancount data for bookkeeping purposes.




```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

To create a scatterplot of the mean values of control vs. treated use this:

```{r}
plot(meancounts)
```


```{r}
hist(meancounts$control.mean, breaks = 300)
```


To make both axes on a logorithmic scale, use the following:

```{r}
plot(meancounts, log= "yx")
```

To add log2foldchange to meancounts data frame

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
```


To get ridd of nonsense results

```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
nrow(mycounts)
```

A common rule of thumb in the field is to use a log2fc of greater than +2 as "upregulated" and less than -2 as "downregulated"

How many of our genes are upregulated upon this drug treatment?

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)


sum(up.ind)
sum(down.ind)
```

##Let's use DESeq2

```{r}
library(DESeq2)
```

To use DESeq, you need to put data into very specific dataframe


```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```


```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

Can organize results by smallest p value and create new subset of p value < 0.05

```{r}
resOrdered <- res[order(res$pvalue),]
res05 <- results(dds, alpha=0.05)
summary(res05)
```

To visualize, can create a volcano plot

```{r}
plot( res$log2FoldChange,  -log(res$padj), 
      xlab="Log2(FoldChange)",
      ylab="-Log(P-value)")
```

Can add cutoff lines for better visualization of useful data

```{r}
plot( res$log2FoldChange,  -log(res$padj), 
 ylab="-Log(P-value)", xlab="Log2(FoldChange)")

# Add some cut-off lines
abline(v=c(-2,2), col="darkgray", lty=2)
abline(h=-log(0.05), col="darkgray", lty=2)
```

