---
title: "Class 17"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Coronavirus

Here we analyze infection data for the 2019 novel Coronavirus COVID-19 (2019-nCoV) epidemic. The raw data is pulled from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus repository.

A CSV file is available here https://github.com/RamiKrispin/coronavirus-csv


```{r}
url <- "https://tinyurl.com/COVID-2019"
virus <- read.csv(url)

tail(virus)
```




> Q1. How many total infected cases are there around the world?

```{r}
total_cases <- sum(virus$cases)
total_cases
```


> Q2. How many deaths linked to infected cases have there been?


Lets have a look at the "$type" column

```{r}
table(virus$type)
```


```{r}
inds <- virus$type == "death"
death_cases <- sum(virus[inds,"cases"])
```




> Q3. What is the overall death rate?

Percent death 

```{r}
round(death_cases/total_cases * 100,2)
```



> Q4. What is the death rate in Mainland China?


```{r}
China_cases <- subset(virus, Country.Region == "Mainland China")

total_China_cases <- sum(China_cases$cases)

China_death <- sum(China_cases[China_cases$type == "death", "cases"])

round(China_death/total_China_cases * 100, 2)
```


> Q5. What is the death rate in Italy, Iran and the US?

Italian Death rate

```{r}
Italy_cases <- subset(virus, Country.Region == "Italy")

total_Italy_cases <- sum(Italy_cases$cases)

Italy_death <- sum(Italy_cases[Italy_cases$type == "death", 
                              "cases"])

round(Italy_death/total_Italy_cases * 100, 2)
```

Iranian Death Rate

```{r}
Iran_cases <- subset(virus, Country.Region == "Iran")

total_Iran_cases <- sum(Iran_cases$cases)

Iran_death <- sum(Iran_cases[Iran_cases$type == "death", "cases"])

round(Iran_death/total_Iran_cases * 100, 2)
```

US Death Rate

```{r}
usa_cases <- subset(virus, Country.Region == "US")

total_usa_cases <- sum(usa_cases$cases)

usa_death <- sum(usa_cases[usa_cases$type == "death", "cases"])

round(usa_death/total_usa_cases *100, 2)
```


